drivers response diversity

### 2024-09-26 03_drivers_response_diversity_Fig3_Fig4.Rmd
figures 4a-c ready. needed for divergence too

# Read in data on response diversity metrics in each lake
resp_div_all <- readRDS("total_models/resp_div_all.rds") |> 
  select(temp:fLake)

@@

##Fig. 3
Dissimilarity, species richness and categories

@@

lm1_plot
# Lms for all categories: endemic, translcated, non_native and non_endemic_native
# with data on all lakes
# Endemic
lm_endemic <- lm(mean_rdiv ~ endemic, data = df_lm)

@@

# save as TIFF
# tiff(paste("total_models/plots/lm_divergence_all.tiff", sep = ""), units="in", width=11, height=7, res=300)
# # plot(ggarrange(depth1, depth2, ncol = 2))
# # plot science discussion
# plot(lm_divergence_all)
# 
# dev.off()
```
## Figure 4a
```{r}
df_lm_4 <- df_lm |> 
  filter(!Lake %in% c("Poschiavo", "Joux", "Maggiore", "Lugano"))

@@

# new data
new_data_species_loss <- tibble(Phos_max_Vonlanthen = seq(from = min(df_lm_4$Phos_max_Vonlanthen), to = max(df_lm_4$Vonlanthen), length = 15))

@@

bind_model_species_loss|> 
  ggplot(aes(Phos_max, fit, color = category, fill = category)) +
  geom_line(linewidth = 1.2) +
  geom_ribbon(aes(ymin = (fit - se.fit), ymax = (fit + se.fit)),  alpha = 0.3, color = NA) +
  theme_classic(base_size = 20) +
  ylab("Species loss") +
  xlab("Maximum total phosphorus (µg/L)") +
  # ylim(1,3) +
  scale_fill_manual(values = mycolor, aesthetics = c("color", "fill"), guide = NULL)

```
## Figure 4b
```{r}
# second lm for maximum phosphorus ~ species richness
lm2 <- lm(mean_rdiv ~ Phos_max + sum_species, data = df_lm_excl)

@@

# checking model assumptions
summary(lm_test)
# shapiro.test(resid(lm_species_loss))
# lmtest::bptest(lm_species_loss)
# plot(lm_species_loss)

>>>>

# Read in data on response diversity metrics in each lake
resp_div_all <- readRDS("total_models/resp_div_all_MA.rds") |> 
  select(temp:fLake)

@@

##Fig. 3a
Dissimilarity, species richness and categories

@@

lm1_plot

##Figure 3b
```{r}
# Lms for all categories: endemic, translcated, non_native and non_endemic_native
# with data on all lakes
# Endemic
lm_endemic <- lm(mean_rdiv ~ endemic, data = df_lm)

@@

# save as TIFF
# tiff(paste("total_models/plots/lm_divergence_all.tiff", sep = ""), units="in", width=11, height=7, res=300)
# # plot(ggarrange(depth1, depth2, ncol = 2))
# # plot science discussion
# plot(lm_divergence_all)
# 
# dev.off()
```
## Figure 4a Log
```{r}

# LOG
df_lm_4 <- df_lm |> 
  filter(!Lake %in% c("Poschiavo", "Joux", "Maggiore", "Lugano"))

# add log 
df_lm_4 <- df_lm_4 |> 
  mutate(P_max_log = log(Phos_max_Vonlanthen))

lm_species_loss_log <- lm(Species_loss ~ P_max_log, data = df_lm_4)

# checking model assumptions
summary(lm_species_loss_log)
shapiro.test(resid(lm_species_loss_log))
lmtest::bptest(lm_species_loss_log)
plot(lm_species_loss_log)

# new data
new_data_species_loss_log <- tibble(P_max_log = seq(from = min(df_lm_4$P_max_log), to = max(df_lm_4$P_max_log), length = 15))

# predict model
prediction_species_loss_log <- predict.lm(lm_species_loss_log, newdata = new_data_species_loss, se.fit = TRUE, type = "response")

# combine new data and predictions
bind_model_species_loss_log <- cbind(new_data_species_loss_log, prediction_species_loss_log) |> 
  mutate(category ="color")

bind_model_species_loss_log|> 
  ggplot(aes(P_max_log, fit, color = category, fill = category)) +
  geom_line(linewidth = 1.2) +
  geom_ribbon(aes(ymin = (fit - se.fit), ymax = (fit + se.fit)),  alpha = 0.3, color = NA) +
  theme_classic(base_size = 20) +
  ylab("Species loss") +
  xlab("Maximum total phosphorus (µg/L)") +
  # ylim(1,3) +
  scale_fill_manual(values = mycolor, aesthetics = c("color", "fill"), guide = NULL)

```
## Figure 4a
```{r}

df_lm_4 <- df_lm |> 
  filter(!Lake %in% c("Poschiavo", "Joux", "Maggiore", "Lugano"))

# add log 
df_lm_4 <- df_lm_4 |> 
  mutate(P_max_log = log(Phos_max_Vonlanthen))

@@

# new data
new_data_species_loss <- tibble(Phos_max_Vonlanthen = seq(from = min(df_lm_4$Phos_max_Vonlanthen), to = max(df_lm_4$Phos_max_Vonlanthen), length = 15))

@@

bind_model_species_loss|> 
  ggplot(aes(Phos_max_Vonlanthen, fit, color = category, fill = category)) +
  geom_line(linewidth = 1.2) +
  geom_ribbon(aes(ymin = (fit - se.fit), ymax = (fit + se.fit)),  alpha = 0.3, color = NA) +
  geom_point(data = df_lm_4, aes(x = Phos_max_Vonlanthen, y = Species_loss), inherit.aes = FALSE, color = "black", size = 3) +
  theme_classic(base_size = 20) +
  ylab("Species loss") +
  xlab("Maximum total phosphorus (µg/L)") +
  # ylim(1,3) +
  scale_fill_manual(values = mycolor, aesthetics = c("color", "fill"), guide = NULL)

## Figure 4b
```{r}
# second lm for maximum phosphorus ~ species richness
lm2 <- lm(mean_rdiv ~ Phos_max + sum_species, data = df_lm_excl)

@@

# checking model assumptions
summary(lm_test)
shapiro.test(resid(lm_test))
lmtest::bptest(lm_test)
plot(lm_test)
